Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-05-09 cmsb22.tex
And more data

In this article we proposed a new method for the computation of minimal trap-spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

@@

\section{Temporary}

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap-spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
  \centering
  \label{tab:pyboolnet_repo}
  \begin{tabular}{lrrrrrr}
    \toprule
    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
    multivalued & 13 &&& 4 & 0.02 & 0.02\\
    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
    raf & 3 &&& 2 & 0.03 & 0.02\\
    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
    \bottomrule
  \end{tabular}

@@

\end{document}

>>>>

In this article we proposed a new method for the computation of minimal trap spaces of Boolean models, based on a new concept called maximal conflict-free siphons.
This method is evaluated on large models from the literature and shows that it can scale up much better than the state-of-the-art prime-implicants based techniques.
We believe that this opens up the way to a much better analysis of large Boolean models, which is needed with the advent of automatic model-generation pipelines~\cite{ostaszewski2021covid19}.

@@

\section{PyBoolNet repo}

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on the PyBoolNet's repository.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
  \centering
  \label{tab:pyboolnet_repo}
  \begin{tabular}{lrrrrrr}
    \toprule
    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
    arellano\_rootstem & 9 &&& 4 & 0.04 & \textbf{0.31}\\
    calzone\_cellfate & 28 &&& 27 & 0.03 & 0.08\\
    dahlhaus\_neuroplastoma & 23 &&& 32 & 0.04 & 0.11\\
    davidich\_yeast & 10 &&& 12 & 0.03 & 0.08\\
    dinwoodie\_life & 15 &&& 7 & 0.03 & 0.08\\
    dinwoodie\_stomatal & 13 &&& 1 & 0.04 & 0.05\\
    faure\_cellcycle & 10 &&& 2 & 0.04 & 0.04\\
    grieco\_mapk & 53 &&& 18 & 0.04 & 0.08\\
    irons\_yeast & 18 &&& 1 & 0.05 & 0.04\\
    jaoude\_thdiff & 103 &&& 1000+ & \textbf{1.73} & 0.40\\
    klamt\_tcr & 40 &&& 8 & 0.06 & 0.04\\
    krumsiek\_myeloid & 11 &&& 6 & 0.03 & 0.03\\
    multivalued & 13 &&& 4 & 0.02 & 0.02\\
    n12c5 & 11 &&& 5 & \textbf{34.37} & 0.06\\
    n3s1c1a & 2 &&& 2 & 0.02 & 0.03\\
    n3s1c1b & 2 &&& 2 & 0.02 & 0.03\\
    n5s3 & 4 &&& 3 & 0.03 & 0.03\\
    n6s1c2 & 5 &&& 3 & 0.03 & 0.04\\
    n7s3 & 6 &&& 3 & 0.02 & 0.02\\
    raf & 3 &&& 2 & 0.03 & 0.02\\
    randomnet\_n15k3 & 15 &&& 3 & 0.04 & 0.05\\
    randomnet\_n7k3 & 7 &&& 10 & 0.03 & \textbf{0.19}\\
    remy\_tumorigenesis & 34 &&& 25 & \textbf{1.79} & 0.13\\
    saadatpour\_guardcell & 13 &&& 1 & 0.02 & 0.03\\
    selvaggio\_emt & 56 &&& 1000+ & \textbf{1.03} & 0.23\\
    tournier\_apoptosis & 12 &&& 3 & 0.05 & 0.06\\
    xiao\_wnt5a & 7 &&& 4 & 0.06 & 0.03\\
    zhang\_tlgl & 60 &&& 156 & 0.26 & 0.17\\
    zhang\_tlgl\_v2 & 60 &&& 258 & 0.19 & 0.17\\
    \bottomrule
  \end{tabular}

@@

\section{Selected models}

\begin{table}[!htb]
  \caption{Timing comparisons between the proposed method and PyBoolNet on selected models from the literature.
    Columns "\(n\)", "\(e\)", and "\(s\)" denote the numbers of nodes, interactions, and source nodes of each model, respectively. Column "\(|M|\)" denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    "DNF" means that the method did not finish the computation (with a maximum of 1000 trap spaces) within the timeout of 2 minutes. Number in bold indicate a ratio greater than three.}
  \centering
  \label{tab:selected_redone}
  \begin{tabular}{lrrrrrr}
    \toprule
    model & $n$ & $e$ & $s$ & $|M|$ & PyBoolNet (s) & Trappist (s) \\ \midrule
    CASCADE3 & 183 &&& 1 & 118.54 & 1.03\\
    KYNURENINE-PATHWAY & 150 &&& 1 & 102.45 & 1.09\\
    SN5 & 2746 &&& ??? & DNF & DNF\\
    ERBB-RECEPTOR-SIGNALING & 247 &&& 1000+ & DNF & 3.56\\
    Executable\_file\_for\_cholocystokinin\_model & 404 &&& 1000+ & 1.46 & 1.79\\
    korkut\_2015a & 99 &&& 1000+ & DNF & 0.56\\
    SIGNALING-PATHWAY-FOR-BUTANOL-PRODUCTION & 66 &&& 1 & 0.24 & 0.16\\
    Corral\_ThIL17diff\_15jan2021 & 92 &&& 1000+ & DNF & 0.32\\
    HIV-1 & 138 &&& 6 & DNF & 0.33\\
    Executable\_file\_for\_CaSQ\_derived\_mast\_cell\_activation\_model & 80 &&& 1000+ & ??? & 0.17\\
    Colitis\_associated\_colon\_cancer & 70 &&& 5 & 0.22 & 0.07\\
    INTERFERON-1 & 121 &&& 2 & 13.21 & 0.12\\
    TLGLSurvival & 61 &&& 3 & 0.73 & 0.16\\
    cho\_2016\_all\_pos & 200 &&& 1000+ & DNF & 0.64\\
    turei\_2016 & 4691 &&& ??? & DNF & DNF\\
    Human\_network & 1953 &&& 1000+ & DNF & 42.58\\
    Executable\_file\_for\_CaSQ\_derived\_MAPK\_model & 182 &&& 1000+ & ??? & 0.32\\
    TCR-TLR5-SIGNALING-2018 & 130 &&& 48 & 1.52 & 0.17\\
    Hernandez\_TcellCheckPoints\_13april2020 & 218 &&& 1000+ & 53.51 & 0.55\\
    IMMUNE-SYSTEM & 164 &&& 1000+ & DNF & 0.92\\
    HMOX1-PATHWAY & 145 &&& 8 & 3.72 & 0.11\\
    Signaling\_in\_Macrophage\_Activation & 321 &&& 1 & 13.25 & 0.31\\
    FIBROBLASTS & 139 &&& 1000+ & DNF & 0.61\\
    S1\_Table & 1659 &&& 1000+ & DNF & 58.98\\
    Regan2020\_Adhesion\_CIP\_Migration\_CellCycle\_Apoptosis & 121 &&& 78 & 53.30 & 0.53\\
    SIGNALING-IN-PROSTATE-CANCER & 133 &&& 1000+ & DNF & 0.46\\
    InflammatoryBowelDisease & 47 &&& 1 & DNF & 33.02\\
    Influenza\_A\_Virus\_Replication\_Cycle & 131 &&& 17 & 51.17 & 0.13\\
    IL\_6\_Signalling & 86 &&& 1 & 0.49 & 0.07\\
    VIRUS-REPLICATION-CYCLE & 154 &&& 2 & DNF & 0.19\\
    Executable\_file\_for\_Alzheimer\_model & 1169 &&& 1000+ & DNF & 1.53\\
    ER-STRESS & 182 &&& 32 & 31.99 & 0.21\\
    RA\_apoptosis\_executable\_module & 180 &&& 1000+ & DNF & 0.23\\
    \bottomrule
  \end{tabular}
\end{table}

\sylvain{Some models seem to have a different number of trap spaces here than in the other table (like RA apoptosis), for others quite large time differences (ERBB, inflammatory bowel)\dots}
\end{document}
